---
title: "Football Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---

```{r}
source("readDataToMemory.R")
readFootball()

library(ggplot2)
library(ggthemes)

print(dbListTables(sql_con))
```

## Pipes

```{r}
players %>%
  head
```

```{r}
players %>%
  colnames
```

```{r}
sapply(players,class)
```
Los mismo podemos hacer con pipes:

Para pasar el primer argumento a la siguiente función:
```{r}
players %>%
  sapply(class)
```

Si queremos pasar el segundo argumento a la siguiente función:
```{r}
class %>%
  sapply(players,.)
```



## SELECT ~ select

```{r}
players %>%
  head %>%
  select(player_name, birthday)
```

```{r}
dbGetQuery(sql_con, "
SELECT player_name
,   birthday
FROM Player
LIMIT 6
")
```

## WHERE ~ filter

```{r}
players %>%
  filter(birthday >= "1998") %>%
  head
```
```{r}
dbGetQuery(sql_con, "
SELECT *
FROM Player
WHERE birthday >= 1998
LIMIT 6
")
```
## ORDER ~ arrange

```{r}
players %>%
  arrange(birthday) %>%
  head
```
```{r}
dbGetQuery(sql_con, "
SELECT *
FROM Player
ORDER BY birthday ASC
LIMIT 6
")
```
## aggregators ~ summarise

```{r}
players %>%
  summarise(weight=mean(weight), height=mean(height))
```
```{r}
dbGetQuery(sql_con, "
SELECT AVG(weight) AS weight
, AVG(height) AS height
FROM Player
")
```

## mutate

```{r}
players %>%
  mutate(weight_kg = weight*0.45359237) %>%
  summarise(weight=mean(weight_kg), height=mean(height))
```
```{r}
dbGetQuery(sql_con, "
SELECT AVG(wkg.weightkg) AS weight
, AVG(height) AS height
FROM (
  SELECT weight*0.45359237 AS weightkg, height
  FROM Player
)wkg
")
```
```{r}
"
SELECT AVG(weight*0.45359237) AS weight
, AVG(height) AS height
FROM Player
" %>%
dbGetQuery(sql_con,.)
  ```

## GROUP BY

```{r}
players %>%
  mutate(year=as.numeric(substring(birthday, 1, 4))) %>%
  group_by(year) %>%
  summarise(height=mean(height)) %>%
  ggplot(aes(year, height)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```
```{r}
"SELECT CAST(SUBSTR(birthday,1,4) AS INT) AS year
  , height
  FROM Player
LIMIT 6
" %>%
dbGetQuery(sql_con,.)
```

```{r}
"SELECT cast(birthday as daytime) AS year
  , height
  FROM Player
LIMIT 6
" %>%
dbGetQuery(sql_con,.)
```

```{r}
"
SELECT AVG(height) AS height, AVG(year) AS year
FROM(
  SELECT cast(birthday as daytime) AS year
  , height
  FROM Player
) new
GROUP BY new.year
" %>%
dbGetQuery(sql_con,.) %>%
  ggplot(aes(year, height)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```

```{r}
"
SELECT AVG(height) AS height, cast(birthday as daytime) AS year
FROM Player
GROUP BY year
" %>%
dbGetQuery(sql_con,.) %>%
  ggplot(aes(year, height)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```
```{r}
"
SELECT CAST(SUBSTR(birthday,1,4) AS INT) AS year
,  COUNT(1) AS n
,  AVG(height) AS height
FROM Player
GROUP BY SUBSTR(birthday,1,4)
" %>%
dbGetQuery(sql_con,.)
```


```{r}
players %>%
  mutate(year=as.numeric(substring(birthday, 1, 4))) %>%
  group_by(year) %>%
  summarise(nsample=n()) %>%
  ggplot(aes(year, nsample)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```
```{r}
"
SELECT COUNT(1) AS nsample, cast(birthday as daytime) AS year
FROM Player
GROUP BY year
" %>%
dbGetQuery(sql_con,.) %>%
  ggplot(aes(year, nsample)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```



```{r}
players %>%
  mutate(year=as.numeric(substring(birthday, 1, 4))) %>%
  group_by(year) %>%
  summarise(height=mean(height), nsamples=n()) %>%
  filter(nsamples >= 200) %>%
  ggplot(aes(year, height)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```
```{r}
"
SELECT COUNT(1) AS nsample, AVG(height) AS height, cast(birthday as daytime) AS year
FROM Player
GROUP BY year
HAVING nsample >=200
" %>%
dbGetQuery(sql_con,.) %>%
  ggplot(aes(year, height)) +
  geom_line(colour="darkorange") +
  theme_minimal()
```


Exercise

* Translate from SQL to dply or from dply to SQL
```{r}
dbGetQuery(sql_con, "
SELECT *
FROM Match
LIMIT 6
")
```

```{r}
dbGetQuery(sql_con, "
SELECT country_id
,   COUNT(1) AS n
FROM Match
GROUP BY country_id
ORDER BY n DESC
")
```
```{r}
countries
```


```{r}
games %>%
  left_join(countries, by=c("country_id" = "id")) %>%
  select(name, home_team_api_id) %>%
  group_by(name) %>%
  summarise(teams = n_distinct(home_team_api_id)) %>%
  arrange(desc(teams))
```

```{r}
dbGetQuery(sql_con, "
SELECT *
FROM Player_Attributes
LIMIT 6
")

```

* From where do players with higest accuracy come from?

* Add team name and translate to dplyr

```{r}
dbGetQuery(sql_con, "
SELECT stage, home_team_api_id, home_team_goal
FROM Match
WHERE country_id = 1
AND season = '2008/2009'
ORDER BY stage
")
```
