---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
---


```{r}
source("readDataToMemory.R")

library(DBI)
library(ggthemes)
library(ggplot2)

readInstacart()


src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
 An example of aggregation:
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}

products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?
* The most popular hour of doing order? 
* The most reordered products?
* The less reorderd product?
* Plot histogram of number of products in an order.
* What is the most popular department? 


* Which day of week people order most fequently?
```{r}  
dbGetQuery(sc, "DESC orders_tbl")

dow_popularity_txt <- "
SELECT order_dow
,   COUNT(1) AS n_orders
FROM orders_tbl
GROUP BY order_dow
ORDER BY n_orders DESC
"

dow_popularity <-
  dbGetQuery(sc, dow_popularity_txt)

dow_popularity
```

* The most popular hour of doing order? 
```{r}  
# dbGetQuery(sc, "DESC orders_tbl")

hour_popularity_txt <- "
SELECT order_hour_of_day
,   COUNT(1) AS n_orders
FROM orders_tbl
GROUP BY order_hour_of_day
ORDER BY n_orders DESC
"

hour_popularity <-
  dbGetQuery(sc, hour_popularity_txt)

hour_popularity
```

* The most reordered products?
```{r}
# dbGetQuery(sc, "DESC order_products__prior_tbl")

products_reorder_popularity_txt <- "
SELECT product_id
,   SUM(reordered) AS n_reorders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_reorders DESC
LIMIT 50
"

products_reorder_popularity <-
  dbGetQuery(sc, products_reorder_popularity_txt)

products_reorder_popularity
```
```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_reorders
,   product_name
FROM (
    SELECT product_id
    ,   SUM(reordered) AS n_reorders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_reorders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

* The less reorderd product?
```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_reorders
,   product_name
FROM (
    SELECT product_id
    ,   SUM(reordered) AS n_reorders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_reorders ASC
    LIMIT 10) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```
```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_reorders
,   product_name
FROM (
    SELECT product_id
    ,   SUM(reordered) AS n_reorders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_reorders ASC
    ) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
WHERE op.n_reorders!=0
LIMIT 10
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```


* Plot histogram of number of products in an order.
```{r}
# dbGetQuery(sc, "DESC order_products__prior_tbl")

# lo hago con order_products__train_tbl porque con 
# order_products__prior_tbl da error de falta de memoria
n_prod_per_order <- dbGetQuery(sc, 
"SELECT order_id
, COUNT (1) AS n_products
FROM order_products__train_tbl
GROUP BY order_id
order by n_products DESC
LIMIT 20
")

n_prod_per_order
```

```{r}
n_prod_per_order <- dbGetQuery(sc, 
"SELECT order_id
, COUNT (1) AS n_products
FROM order_products__train_tbl
GROUP BY order_id
")

hist(n_prod_per_order$n_products,br=80)
```

```{r}
  ggplot(n_prod_per_order,
    aes(n_products)) +
  geom_hist( fill='darkorange') +
  theme_minimal()
```


* What is the most popular department? 


## Whisker, magrittr

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).


* For given day, rank department popularity.

## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```

